|

### 2020-11-16 paper/v1/broadband.speed.covid.Rmd
merged intros

**PARA1**During the pandemic, working from home using Information and Communication Technologies (ICT), whether partially or exclusively, was transformed from a niche means of accessing work, albeit one that had been on a slow, upward trend, to a widespread way of life in many countries (Felstead and Reuschke, 2020; **insert further refs from elsewhere**). The ability to work from home or telecommute has enabled millions to retain their jobs and, to a varying extent, maintain productivity during periods of strict lockdown around the world. However, this ability has not been evenly distributed socially or spatially, creating a new type of digital divide. On one side are those who can work from home and have been able to enjoy both economic resilience and greater personal safety. On the other side, previously employed individuals have been forced to accept furlough or redundancy packages unless they are part of the cadre of essential workers, who are potentially at high risk of infection. Using the UK as a case study, this paper aims to improve our understanding of the spatial and social dimensions of this new digital divide by assessing the resilience of broadband speeds in terms of both quality and reliability of service, and whether this reinforces or redresses prior digital divisions.

HB
**PARA2**The capability to work from home has previously been studied from the perspective of whether work tasks in a given occupation both can be and are allowed to be performed independently of location or co-location with colleagues, including supervisors (**insert refs**). However, successful telecommuting also requires that the quality and reliability of ICT services, particularly home internet connection speeds, enable the completion of work tasks with a minimum of delay or interruption. In reality, the performance of broadband speeds with respect to telecommuters has never been tested before at scale, as working from home and connecting to colleagues and workplace resources via broadband has previously been the purview of a small minority of workers. In the UK, the most recent statistics indicate that **x%** of working age adults identified their home as their usual workplace, with an additional **x%** who said they work from home at least once a week (**insert stats from 2018 or 2019 NTS - also x% of these use ICT to complete their work at home**). Their demands on internet services have been minor compared to the demands of leisure users, such that the broadband performance offered by different Internet Service Providers (ISPs) has been benchmarked according to download speeds during the evening 'primetime', when video streaming services are at their peak, rather than during the working  day (**OfCom, 2017**). Yet the pandemic has fundamentally changed not only how many people work from home - 47% of working age adults at the peak in April, 2020 (**ref**) - but also their technical requirements when doing so. Whilst it may have been common in the past for telecommuters to complete solitary work tasks at home, but still visit their workplace for meetings (**ref?**), during the pandemic the replacement of face-to-face contact with bandwidth-intensive video conference calling was seen as essential to suppressing the spread of infection. This type of broadband use can particularly affect upload speeds **ref?**, so we seek to develop a new benchmark (**is it a benchmark?**)to assess the quality and reliability of upload speeds by time of day and day of the week across local authorities in the UK during the time when the population were told to work from home if at all possible.

**PARA3**The experience of upload speeds is not the same as the maximum speeds offered by an ISP or possible speeds over a particular type and length of connection. Therefore, we use volunteered geographic information in the form of speed-checks run by users to test their experienced broadband speeds, upload and download. Meanwhile, the quality and reliability of upload speeds, like ICT more generally, vary in time and space due to both supply and demand-side influences, and can be measured in a number of ways. These include: a) mean, experienced connection speed, b) standard deviation or the amount of fluctuation from the mean, and c) the variation in speeds at particular times of day when working from home is more likely to take place. We take account of all three measurements in order to describe upload speeds as fully as possible. We start by calculating the mean upload speed for every local government district in the UK for each hour of each weekday, excluding midnight to 6:00, weekends and bank holidays. We then cluster the local authorities by these temporal profiles, allowing us to identify patterns and describe the overall means, standard deviations, and other relevant statistics for each spatial cluster. Thus we aim to answer the first part of our research question: How resilient are broadband speeds as experienced in different parts of the UK during a time of extreme demand?

**PARA4**The cause of these different experiences of broadband resilience may be different in different areas, as they may reflect similarities in patterns of demand or similar quality of infrastructure. Our approach is also limited by potential endogeneity, as for example, better quality connections with high mean speeds may enable more working from home, but greater demand may cause slower speeds, less reliability and greater variability of speed at different times of day or week. Therefore, we avoid attributing any cause to our analysis of the experienced level of quality and reliability of upload speeds. Instead, we run auxilliary regressions in order to understand how the spatial and temporal patterns of internet service relate to the economic geography of the UK. We discuss how the different patterns might support or undermine efforts to work from home and maintain safe productivity and whether they reinforce existing spatial and social inequalities. From this analysis, we hope to provide a greater understanding of how telecommuting and technology intersect at a time of extreme demand, and what lessons this time has for a future where telecommuting is likely to remain a common means of accessing work and broadband services, as well as infrastructure, must be fit for purpose.

**PARA5*The structure of this paper is as follows.

## broadband studies, digital divides, broadband tech stuff - benchmarking?

## from telecommuting to #WFH - Introduce space-time geography - circumstances where work can be carried out more flexibly in space and time- how relate to digital divides - capability theory of mobility? suppressed demand...

##covid and working from home, cities, urban structure - economic geography of UK
Potentially useful readings for covid and cities:

- https://www.coronavirusandtheeconomy.com/question/why-has-coronavirus-affected-cities-more-rural-areas
- https://www.coronavirusandtheeconomy.com/question/will-coronavirus-cause-big-city-exodus
- [EPB commentaries](https://journals.sagepub.com/toc/EPB/current)
- https://journals.sagepub.com/doi/full/10.1177/2399808320926912?casa_token=gh3SgKFCZ44AAAAA:MjjMWJ61DqlmxzkWQyy_wxPnU20QdyQYkC4fVFpHLPfLk7McwmQGkJ7x2Q7LZOjTf6vcYaqNwug
- [WFM and covid, econ](https://www.coronavirusandtheeconomy.com/question/what-has-coronavirus-taught-us-about-working-home)
- [productivity and WFM](https://www.coronavirusandtheeconomy.com/question/who-can-work-home-and-how-does-it-affect-their-productivity)

ET
During the pandemic, working from home using Information and Communication
Technologies (ICT), whether partially or exclusively, was transformed from a
niche means of accessing work, albeit one that had been on a slow, upward trend,
to a widespread way of life in many countries.
The ability to work from home or telecommute meant millions retained their jobs
and, to a varying extent, maintained productivity during periods of strict
lockdown around the world.
However, this ability was not evenly distributed socially or spatially, creating
a new type of digital divide -- one of economic resilience and personal safety on
one side and unemployment or essential front-line work on the other.
The ability to telecommute is partly a product of whether the majority of work
tasks in an occupation can be performed independently of being in a particular
location or co-locating with colleagues.
It is also a product of the quality of ICT services, particularly home internet
connections, sufficient to successfully complete work tasks with a minimum of
delay or interruption.
Therefore, in this paper, we look at patterns of broadband speed tests
and upload speeds in the UK in order to gain insights into both the quality and
reliability of broadband services across the country and the distribution of
telecommuting at the time when the population were told to work from home if
possible.
To do so, we employ unique data regarding individual broadband speed tests and
state of the art time series clustering methods.
Our analysis enables us to better understand how telecommuting and
technology intersect at a time of extreme demand, and what lessons this time has
for a future where telecommuting is likely to remain a common means of accessing
work and broadband services, as well as infrastructure, must be fit for purpose.
**I THINK WE NEED A MORE CLEAR RESEARCH QUESTION**

@@

<!-- Who works from home-->
There is a consensus that opportunities for working from home especially during
the current pandemic are not equally spread across the workforce.
@NBERw26948 indicated that in the US managers, educators, as well as those
working in computer-related occupations, finance, and law can easily work from
home.
On the contrary, workers in farming, construction, and manufacturing do not have
such opportunities.
Not surprisingly, occupations with opportunities to telecommute are associated
with higher earnings.
This is not the case for the workforce occupied in less footloose occupations as
as they tend to be lower-income, non-white, without a university degree,
live in rental accommodation and lack health insurance [@NBERw27085].
Although these figures refer to the US, similar trends can be observed for other
countries.
For example, $75$% of workers with tertiary education worked from home in Europe
during spring $2020$, while the same share for workers with secondary and
primary education dropped to $34$% and $14$% respectively.
Moreover, employees living in cities, women and younger employees were
have worked from home [@eurofound2020].

<!-- ICT infrastructure-->
None of these changes could have happened in the absence of ICT
infrastructure -- both in terms of software and hardware.
But while innovations in software are easily diffused across space and
society^[See for example the huge success of videoconferencing apps
[@marks2020zoom].],
the same does not apply for ICT hardware infrastructure such as internet
broadband connectivity.
The literature exemplifies digital divides in terms of internet access and its
quality.
For instance, @riddlesden2014broadband highlighted the broadband divides in the
UK, while the systematic review from @SALEMINK2017360 reinforced our
understanding for the infrastructure quality differences between urban and rural
areas.

<!-- WFH and digital divides-->
Our framework to understand telecommuting as a function of occupations and
quality of ICT infrastructure is aligned with current debates on digital divides.
While the, so-called, first level digital divides are associated with access
and quality of internet connectivity, the second level ones are linked
to the necessary skills to effectively utilise ICT and the internet
[@blank2014dimensions; @van2011internet].
Importantly, the differentiated capacity to telecommute, which to a certain
extend are related to the first and second level digital divides, leads to
differentiated outcomes regarding the economic resilience of people and
places against the current pandemic.
And to the extend that the quality of the internet infrastructure and occupation
variation are spatially dependent and clustered in space, the spatial footprint
of telecommuting is of great interest. In a way similar to the third level
digital divide, which
focuses on the differentiated returns of internet use [@stern2009levels;
@van2014digital; @van2015third] places with high rates of telecommuting during
the Covid pandemic illustrate higher economic resilience against the current pandemic.
<!-- Econ resilience-->
This is aligned with the regional economic resilience literature, which
underlines the differentiated capacity of cities and regions to escape or
recover from economic crises [@martin2012regional; @kitsos2018economic].

<!-- WFH broader consequences-->
The long-term effects of such drastic changes in telecommuting and attitudes
towards working from home are difficult to predict. Nevertheless,
they might span through various aspects of economy and society: from changes to
transportation planning due to changes in commuting patters to changes in land
use and urban planning to accommodate people who work from home
(@BUDNITZ2020102713); and from productivity and innovation changes to a change
in agglomeration externalities and the attraction of large cities [@econobs]
just to name a few. This paper is positioned to support such endeavours by
providing an explanatory **[NOTE SURE ABOUT THIS]** framework  based on
the quality of internet connectivity and the frequencies of occupations within
Local Authorities in the UK.

@@

**PARA5**: Data and methods

**PARA6**: Contribution:

@@

@@

@@

In this analysis, the terms ‘teleworking’ and ‘working from home’ are used interchangeably. An appropriate definition of teleworking is ‘the remote
provision of labour that would otherwise be carried out within company premises’ (European Commission, 2020b). In practice, during the COVID crisis,
most such work was carried out in the homes of individual employees rather than any other location.

**PARA about Contention**

Desription of the method

@@

@@

@@

##Upload Clusters / cluster description

@@

##aux regressions

>>>>

- [our google doc](https://docs.google.com/document/d/1PWjkmgzWGYKR9wFogKYw7l-8mZLoORt593x-Tu-f2-M/edit#heading=h.i5om1o8wpcd9)

During the pandemic, working from home using Information and Communication
Technologies (ICT), whether partially or exclusively, was transformed from a
niche means of accessing work, albeit one that had been on a slow, upward trend,
to a widespread way of life in many countries [@felstead2020homeworking].
**HANNAH, DO WE REALLY NEED A REFERENCE IN THE PREVIOUS SENTENCE?**
The ability to work from home or telecommute meant millions retained their jobs
and, to a varying extent, maintained productivity during periods of strict
lockdown around the world.
<!--OLD TEXT: However, this ability was not evenly distributed socially or spatially, creating
a new type of digital divide -- one of economic resilience and personal safety on
one side and unemployment or essential front-line work on the other.
The ability to telecommute is partly a product of whether the majority of work
tasks in an occupation can be performed independently of being in a particular
location or co-locating with colleagues.-->
However, this ability has not been evenly distributed socially or spatially,
creating a new type of digital divide.
On one side are those who can work from home and have been able to enjoy both
economic resilience and greater personal safety.
On the other side, previously employed individuals have been forced to accept
furlough or redundancy packages unless they are part of the cadre of essential
workers, who are potentially at high risk of infection.
Using the UK as a case study, this paper aims to improve our understanding of
the spatial and social dimensions of this new digital divide by assessing the
resilience of broadband speeds in terms of both quality and reliability of
service, and whether this reinforces or redresses prior digital divisions.
<!--and upload speeds in the UK in order to gain insights into both the quality and
reliability of broadband services across the country and the distribution of
telecommuting at the time when the population were told to work from home if
possible.-->
To do so, we employ unique volunteered geographic data on individual broadband
speed tests and state-of-the-art time-series clustering methods, which enable us
to create clusters of UK local authorities with similar temporal signatures of
experienced internet speeds.
We then associate these clusters of local authorities with their socioeconomic
and geographic characteristics to explore whether they are linked with existing
divides and the economic geography of the UK.
**HANNAH, HOW DOES THIS SOUND?**
Our analysis enables us to better understand how telecommuting and
technology intersect at a time of extreme demand, and what lessons this time has
for a future where telecommuting is likely to remain a common means of accessing
work and broadband services, as well as infrastructure, must be fit for purpose.
**LET'S LEAVE IT FOR NOW, BUT I THINK WE CAN CRYSTALISE MORE THE RQ**

<!-- Our framework: capability to WFH; HB old PARA2 -->
The capability to work from home has previously been studied from the
perspective of whether work tasks in a given occupation both can be and are
allowed to be performed independently of location or co-location with
colleagues, including supervisors (**insert refs**).
However, successful telecommuting also requires that the quality and
reliability of ICT services, particularly home internet connection speeds,
enable the completion of work tasks with a minimum of delay or interruption.
In reality, the performance of broadband speeds with respect to telecommuters
has never been tested before at scale, as working from home and connecting to
colleagues and workplace resources via broadband has previously been the
purview of a small minority of workers.
This papers understands telecommuting as a function of the quality of ICT
services, particularly home internet connections, and whether work
tasks in an occupation can be performed independently of being in a particular
location or co-locating with colleagues

@@

<!-- Who works from home-->
There is a consensus that opportunities for working from home especially during
the current pandemic are not equally spread across the workforce.
@NBERw26948 indicated that in the US managers, educators, as well as those
working in computer-related occupations, finance, and law can easily work from
home.
On the contrary, workers in farming, construction, and manufacturing do not have
such opportunities.
Not surprisingly, occupations with opportunities to telecommute are associated
with higher earnings.
This is not the case for the workforce occupied in less footloose occupations as
they tend to be lower-income, non-white, without a university degree,
live in rental accommodation and lack health insurance [@NBERw27085].
Although these figures refer to the US, similar trends can be observed for other
countries.
For example, $75$% of workers with tertiary education worked from home in Europe
during spring $2020$, while the same share for workers with secondary and
primary education dropped to $34$% and $14$% respectively.
Moreover, employees living in cities, women and younger employees were
have worked from home [@eurofound2020].

<!-- ICT infrastructure-->
None of these changes could have happened in the absence of reliable ICT
infrastructure -- both in terms of software and hardware.
But while software innovations are easily diffused across space and
society^[See for example the huge success of videoconferencing apps
[@marks2020zoom].],
the same does not apply for ICT hardware infrastructure such as internet
broadband connectivity.
The literature exemplifies digital divides in terms of internet access and its
quality.
For instance, @riddlesden2014broadband highlighted the broadband divides in the
UK, while the systematic review from @SALEMINK2017360 reinforced our
understanding for the infrastructure quality differences between urban and rural
areas.

<!-- WFH and digital divides-->
Our framework to understand telecommuting as a function of occupations and
quality of ICT infrastructure is aligned with current debates on digital divides.
While the, so-called, first level digital divides are associated with access
and quality of internet connectivity, the second level ones are linked
to the necessary skills to effectively utilise ICT and the internet
[@blank2014dimensions; @van2011internet].
Importantly, the capacity to telecommute, which to a certain
extend is related to the first and second level digital divides, leads to
differentiated outcomes regarding the economic resilience of people and
places against the current pandemic.
And to the extend that the quality of the internet infrastructure and occupation
variation are spatially dependent and clustered in space, the spatial footprint
of telecommuting is of great interest. In a way similar to the third level
digital divide, which
focuses on the differentiated returns of internet use [@stern2009levels;
@van2014digital; @van2015third] places with high rates of telecommuting during
the Covid pandemic illustrate higher economic resilience against the current pandemic.
<!-- Econ resilience-->
This is aligned with the regional economic resilience literature, which
underlines the differentiated capacity of cities and regions to escape or
recover from economic crises [@martin2012regional; @kitsos2018economic].

<!-- WFH broader consequences-->
The long-term effects of such drastic changes in telecommuting and attitudes
towards working from home are difficult to predict. Nevertheless,
they span through various aspects of economy and society: from changes to
transportation planning due to altered commuting patters, to changes in land
use and urban planning to accommodate people who work from home
[@BUDNITZ2020102713]; and from productivity and innovation changes, to changes
in agglomeration externalities and the attraction of large cities [@econobs]
just to name a few. This paper is positioned to support endeavours in
understanding the effects of increased telecommuting by
exposing the spatial and social dimensions of telecommuting including the
resilience of broadband speeds in terms of both quality and reliability of
service, and whether this reinforces or redresses prior digital divisions.
**HANNAH, HOW DOES THIS SOUND**

The structure of this paper is as follows.
...

@@

@@

- PARA about Contention??

<!-- WFH and infrastructure, HB old PARA2 -->
The infrastructural demands of working from home on internet services have been
minor compared to the demands of leisure users, such that the broadband
performance offered by different Internet Service Providers (ISPs) has been
benchmarked according to download speeds during the evening 'primetime',
when video streaming services are at their peak, rather than during the working
day (**OfCom, 2017**).
Yet the pandemic has fundamentally changed not only how many people work from
home, but also their technical requirements when doing so.
Whilst it may have been common in the past for telecommuters to complete
solitary work tasks at home, but still visit their workplace for meetings
(**ref?**), during the pandemic the replacement of face-to-face contact with
bandwidth-intensive video conference calling was seen as essential to
suppressing the spread of infection.
This type of broadband use can particularly affect upload
speeds **ref?**.
This paper offers a framework to assess the quality and reliability of upload
speeds by time of day and day of the week across local authorities in the UK
during the time when the population were told to work from home if at all
possible.

@@

- Introduce space-time geography
- circumstances where work can be carried out more flexibly in space and time
- how relate to digital divides
- capability theory of mobility? suppressed demand...

@@

In this analysis, the terms ‘teleworking’ and ‘working from home’ are used interchangeably. An appropriate definition of teleworking is ‘the remote
provision of labour that would otherwise be carried out within company premises’ (European Commission, 2020b). In practice, during the COVID crisis,
most such work was carried out in the homes of individual employees rather than any other location.
[@eurofound2020]

## covid and working from home, cities, urban structure
- economic geography of UK

@@

<!-- HB old PARA3 -->
The experience of upload speeds is not the same as the maximum speeds offered by an ISP or possible speeds over a particular type and length of connection. Therefore, we use volunteered geographic information in the form of speed-checks run by users to test their experienced broadband speeds, upload and download. Meanwhile, the quality and reliability of upload speeds, like ICT more generally, vary in time and space due to both supply and demand-side influences, and can be measured in a number of ways. These include: a) mean, experienced connection speed, b) standard deviation or the amount of fluctuation from the mean, and c) the variation in speeds at particular times of day when working from home is more likely to take place. We take account of all three measurements in order to describe upload speeds as fully as possible. We start by calculating the mean upload speed for every local government district in the UK for each hour of each weekday, excluding midnight to 6:00, weekends and bank holidays. We then cluster the local authorities by these temporal profiles, allowing us to identify patterns and describe the overall means, standard deviations, and other relevant statistics for each spatial cluster. Thus we aim to answer the first part of our research question: How resilient are broadband speeds as experienced in different parts of the UK during a time of extreme demand?

<!-- HB old PARA4 -->
The cause of these different experiences of broadband resilience may be different in different areas, as they may reflect similarities in patterns of demand or similar quality of infrastructure. Our approach is also limited by potential endogeneity, as for example, better quality connections with high mean speeds may enable more working from home, but greater demand may cause slower speeds, less reliability and greater variability of speed at different times of day or week. Therefore, we avoid attributing any cause to our analysis of the experienced level of quality and reliability of upload speeds. Instead, we run auxilliary regressions in order to understand how the spatial and temporal patterns of internet service relate to the economic geography of the UK. We discuss how the different patterns might support or undermine efforts to work from home and maintain safe productivity and whether they reinforce existing spatial and social inequalities. From this analysis, we hope to provide a greater understanding of how telecommuting and technology intersect at a time of extreme demand, and what lessons this time has for a future where telecommuting is likely to remain a common means of accessing work and broadband services, as well as infrastructure, must be fit for purpose.

@@

Description of the method

@@

## Upload Clusters / cluster description

@@

## aux regressions
